Nonlinear Model Predictive Control-Based Local Planner with Integrated Dynamic Obstacle Avoidance for Mobile Robots

### 2025-04-25 src/amr_control/docs/paper.tex
Added docker compose file, modified CMakellists for amr_control package

\documentclass[conference]{IEEEtran}

\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{fancyhdr}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{subcaption}
\usepackage{blindtext}
\usepackage{microtype}
\usepackage{booktabs}
\usepackage{caption}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{algorithm}
\usepackage{algpseudocode}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em

    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

% Create custom colour for text highlighting
\definecolor{twblue}{HTML}{1f497d}
\definecolor{structure}{HTML}{1f497d}
\definecolor{blindtext}{HTML}{D3D3D3}

\pagestyle{fancy}
\fancyhf{}
\rfoot{\thepage} % Seitenzahl im rechten Fuß

\title{nMPC Local Planner (ändert sich noch}

\author{\IEEEauthorblockN{\textbf{Student:} Eppacher Kevin, BSc. \textbf{PK:} 2310331013}
        \IEEEauthorblockN{\textbf{Peer ReviewerIn:} Schweiger Simon, MSc.}
}

@@

\begin{abstract}
    Mobile robots are widely used in industries, households, and public spaces, such as Autonomous
    Mobile Robots (AMRs) for transporting goods in intralogistics, vacuum/mower robots, and service
    robots in restaurants. However, mobile robots face several challenges, including highly dynamic
    environments, external disturbances such as slippage, which can complicate localization over 
    long distances, and other environmental factors. The goal of this work is to develop a system that 
    enables a mobile robot to navigate optimally and collision-free. Several established approaches,
    such as the Dynamic Window Approach (DWA) and Move Base Flex, have already proven effective.
    The Dynamic Window Approach limits the robot's velocity space to ensure collision avoidance, while
    Move Base Flex provides a flexible interface for path planning and execution, allowing for the 
    integration of different navigation frameworks. In this work, a control system is developed for 
    dynamic obstacle avoidance using a nonlinear Model Predictive Controller (nMPC). The results are 
    compared with the well-established Move Base DWA from ROS1. Both controllers are evaluated in a 
    Gazebo simulation under three conditions: with static objects, with dynamically detected obstacles, 
    and without obstacle avoidance (i.e., without laser scan).

    (Results will follow...)
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
% Introduction to the paper’s main subject (subject = lokale Pfadplanung)

Mobile robots are becoming increasingly indispensable in various fields such as intralogistics, public, and private services, with applications ranging from autonomous floor-cleaning robots to vacuum cleaners. In dynamic environments, these robots are frequently confronted with fast-moving obstacles, which they must detect and avoid in real-time. The conventional approach to navigation involves the mobile robot first creating a map of its surroundings, typically using an occupancy grid to store information about static objects such as walls. Based on this map, a global path is planned using algorithms such as Rapidly Exploring Random Tree (RRT), which the robot subsequently follows \cite{thrun2005probabilistic}.

However, it is important to note that while the global path provides an initial route, it is updated at a lower frequency compared to the local planner, which operates at a higher update rate. The local planner is responsible for adjusting the robot's trajectory in real-time to avoid newly detected obstacles, while still adhering to the waypoints of the global path \cite{siegwart2011autonomous}. This separation between global path planning and local obstacle avoidance is critical in ensuring that the robot can navigate effectively in both known and dynamic environments \cite{khatib1986real}.

@@

\cite{ASAP} conducted a comprehensive comparison of its developed local planner, ASAP, with other seminal planners, including DWA, TEB, and MPC. In Figure \ref{fig:DWA_TEB_MPC_ASAP_path}, the performance of these planners is compared as they guide a mobile robot to follow a global path while avoiding obstacles.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1\columnwidth]{images/DWA_TEB_MPC_ASAP_path.png}
    \caption{Blockdiagram of nMPC for future Implementation }
    \label{fig:DWA_TEB_MPC_ASAP_path}
\end{figure}

In this study, \cite{ASAP} evaluated the local planners based on metrics such as arrival time, cross track error, and computation time.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1\columnwidth]{images/DWA_TEB_MPC_ASAP_stats.png}
    \caption{Blockdiagram of nMPC for future Implementation }
    \label{fig:DWA_TEB_MPC_ASAP_stats}
\end{figure}

@@

Similarly, \cite{MPCvsTEB} performed a comparison between MPC and TEB planners. The study evaluated both the TEB planner and various forms of MPC in terms of arrival time, path length, control effort, and CPU time. The results indicated that the TEB planner and the quadratic form of MPC had comparable arrival times, around 126.9 seconds. However, the time-optimal MPC, which is specifically optimized for minimizing arrival time, completed the task in just 116 seconds. In terms of path length and control effort, both planners showed similar performance. However, the computational time for MPC was found to be three times higher than that of the TEB planner, demonstrating the trade-off between efficiency and computational cost.

% Problem description arising from motivation (was fehlt im aktuellen sota -> vermeidung dynamischer Objekte)
\cite{ASAP} and \cite{MPCvsTEB} illustrate promising results for the MPC as local planner, especially due to the high adaptability forming the cost function for the required needs and the model adaptability, as for example omni-drive-directional-wheels. However, it is shown, that the MPC planner in combination with dynamic obstacle avoidance lack in computation time.

% Paper contribution
The main objective is to design a controller that effectively addresses the challenges of dynamic obstacle avoidance, offering improved performance in real-time navigation within dynamic environments. This work contributes to the state-of-the-art development, of a nonlinear Model Predictive Controller for a mobile robot with integrated dynamic obstacle avoidance. This work proposed a nMPC local Planner, in which the nMPC is integrated with a obstacle avoidance algorithm, with the aim of reducing computation time and increasing the performance, which is evaluated by reducing the arrival time and the cross tracking error.

@@

In this section, we provide an overview of several state-of-the-art local planners, including the Dynamic Window Approach (DWA), Timed Elastic Band (TEB), Model Predictive Control (MPC), Move Base Flex (MBF), Agile and Safe Pursuit (ASAP), and Reinforcement Learning (RL)-based planners. We explain how these planners function and highlight their key limitations, especially when operating in dynamic environments. Additionally, we discuss how nonlinear Model Predictive Control (nMPC) addresses these limitations and provides improved performance.

\subsection{Dynamic Window Approach (DWA)}
DWA is a popular local planner that selects safe and feasible velocities by sampling translational \((v)\) and rotational \((\omega)\) velocities within a dynamic window. These sampled velocities are evaluated through an objective function that balances forward movement and obstacle avoidance. DWA computes a set of trajectories and selects the one that maximizes progress toward the goal while ensuring the robot avoids collisions.

The velocities \(v\) and \(\omega\) are constrained within bounds, as shown in Equation \eqref{eq:dwa_constraints}:

@@

The objective function \(G(v, \omega)\) evaluates the alignment with the goal, distance to obstacles, and forward velocity, as described in Equation \eqref{eq:dwa_objective}:

@@

While DWA is efficient and handles real-time obstacle avoidance well, it struggles with dynamic obstacles due to its lack of future obstacle movement prediction. The extensive velocity sampling can lead to increased computation times in complex environments.

@@

\subsection{Timed Elastic Band (TEB)}
TEB models the robot’s trajectory as a dynamic band that adjusts to avoid obstacles while satisfying kinodynamic constraints. The trajectory is continuously optimized using a cost function that takes into account the goal, collisions, and system dynamics.

@@

\begin{equation}
\label{eq:teb_cost}
J(\mathbf{x}, \mathbf{v}) = \sum_{i=1}^{n-1} \left( \text{cost}_{\text{goal}}(x_i) + \text{cost}_{\text{collision}}(x_i) + \text{cost}_{\text{kinodynamics}}(v_i) \right)
\end{equation}

@@

Recent improvements to TEB incorporate dynamic obstacle velocities using Kalman filtering to predict obstacle movement and optimize the trajectory in real time. This improved TEB method \cite{chen2021dynamic} enhances real-time performance by reducing the overall computational burden while maintaining a high success rate for dynamic obstacle avoidance.

\subsection{Model Predictive Control (MPC)}
MPC computes optimal control inputs over a prediction horizon by solving an optimization problem that accounts for both system dynamics and environmental constraints. Given the system model \(x_{k+1} = f(x_k, u_k)\), MPC minimizes the cost function defined in Equation \eqref{eq:mpc_cost}:

\begin{equation}
\min_{\vec{u}_k} J(\vec{x}_k, \vec{u}_k) = \phi(\vec{x}(N)) + \sum_{k=0}^{N-1} L(\vec{x}(k), \vec{u}(k))
\end{equation}

MPC offers smooth, optimized control over nonlinear dynamics, making it ideal for complex environments. However, it is computationally demanding, especially in real-time applications. Additionally, MPC requires highly accurate models of the robot and environment, which may not always be feasible in dynamic situations.

Compared to DWA and TEB, MPC offers improved handling of dynamic obstacles by continuously adjusting the robot's trajectory in real-time, optimizing for both safety and performance. It also handles nonlinear dynamics more effectively, making it a robust solution for highly dynamic environments.

@@

This section describes the methodology applied in this work, focusing on the integration of the trajectory planner, nMPC, and the Obstacle Detection Algorithm, culminating in the nMPC local planner. Figure \ref{fig:nMPC_Local_Planner_Diagram_2} illustrates the individual components of the nMPC local planner. The main elements are the trajectory planner, nMPC, and optionally, the obstacle detection algorithm, which will be explained in more detail later.

The trajectory planner receives the global path from MoveBase. It processes the global plan into a local plan, taking into account the number of predictions and the prediction horizon length specified by the nMPC. After the nMPC receives the reference trajectory, the robot’s current pose, and the target state, the optimizer of the nMPC calculates the optimal control input and sends it back to the trajectory planner. Depending on whether the differential drive mobile robot (DDMR) has reached its target or not, the trajectory planner then sends the optimal control input \(\mathbf{u} = \begin{bmatrix} v & \omega \end{bmatrix}^\top\)

The robot's absolute position is determined using AMCL, which helps eliminate cumulative slip and drift errors. However, a map must be created in advance. Based on the scan data from the LiDAR, the DDMR can detect obstacles. The Obstacle Detection Node filters the LiDAR data to define a search cone, reducing unnecessary computational load on the nMPC. It then sends the positions of the obstacles, along with the predefined safety radius of the obstacles, to the trajectory planner, which passes this information to the nMPC for further processing.

@@

\subsection{nMPC}

The nMPC is responsible for planning a path based on the current pose, the target pose, and the reference trajectory, using an optimizer that not only satisfies the kinematics of a differential drive mobile robot (DDMR) but also avoids obstacles and respects limitations on control inputs and states.Figure \ref{fig:MPC_Blockdiagram_Implementation} illustrates the components of an MPC.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1\columnwidth]{images/MPC_Blockdiagram_Implementation.png}
    \caption{Detailed Blockdiagram of nMPC \cite{Effects_of_roll_dynamics_for_car_stability_control_by_laguerre_functions}}
    \label{fig:MPC_Blockdiagram_Implementation}
\end{figure}

First, the Optimal Control Problem (OCP) must be converted into a Nonlinear Programming Problem (NLP) using the multiple shooting discretization method. The goal is to minimize the cost function in Equation \ref{eq:L_function}, which consists of the states and control actions for all predictions.

\begin{equation}
\label{eq:L_function}
\begin{aligned}
\min_{\vec{u}_k} J(\vec{x}_k, \vec{u}_k) = J_{\text{obs}} + \phi(\vec{x}(N)) 
+ &\\ \sum_{k}^{k+N-1}
(\vec{x}_{R,k} - \vec{x}_{\text{ref},k})^T Q (\vec{x}_{R,k} - \vec{x}_{\text{ref},k})&+ \vec{u}_k^T R \vec{u}_k
\end{aligned}
\end{equation}

\begin{align}
\text{subject to: } \quad \vec{x}(k+1) &= f(\vec{x}(k), \vec{u}(k)), \\
\vec{x}(0) &= \vec{x}_0, \\
\vec{u}(k) &\in U, \quad \forall k \in [0, N-1], \\
\vec{x}(k) &\in X, \quad \forall k \in [0, N].
\end{align}

The state vector of the robot is described by Equation \ref{eq:state_vector} , and the control vector is described by Equation \ref{eq:control_vector}.

\begin{equation}\label{eq:state_vector}
    \mathbf{x} = \begin{bmatrix} x & y & \theta \end{bmatrix}^T
\end{equation}

\begin{equation}\label{eq:control_vector}
    u = \begin{bmatrix}
    v & \omega
\end{bmatrix}^T
\end{equation}

The equality constraints of the cost function ensure that the robot adheres to its dynamics. The inequality constraints enforce the limits on control inputs and guarantee that the robot avoids obstacles, as expressed in Equation \ref{eq:collision_condition}.

\begin{equation}
\label{eq:phi_definition}
\phi(\vec{x}(N)) = \frac{1}{2} \vec{x}_N^T S \vec{x}_N
\end{equation}

\begin{equation}
J_{\text{obs}} = \sum_{k=0}^{N} \sum_{i=1}^{n_{\text{obs}}} \frac{w_{\text{pen}}}{\sqrt{(x_k - x_{\text{obs},i})^2 + (y_k - y_{\text{obs},i})^2} + \epsilon}
\end{equation}

\begin{equation}
\label{eq:u_constraints}
\vec{u}_{\text{min}} \leq \vec{u}_k \leq \vec{u}_{\text{max}}, \quad \forall k \quad \text{…} \quad \vec{u} \in \mathbb{R}^2
\end{equation}

\begin{equation} \label{eq:1}
    x_{k+1} = Ax_k + Bu_k
\end{equation}

\begin{equation} \label{eq:discrete_state_space_equation_DDMR}
\begin{bmatrix}
    x_{k+1} \\
    y_{k+1} \\
    \theta_{k+1}
\end{bmatrix}
=
\begin{bmatrix}
    x_k \\
    y_k \\
    \theta_k
\end{bmatrix}
+ \Delta T
\begin{bmatrix}
    \cos(\theta) & 0 \\
    \sin(\theta) & 0 \\
    0 & 1
\end{bmatrix}
u_k
\end{equation}

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Verwendete Packages
\usepackage[utf8]{inputenc} % Zeichen-Enkodierung
\usepackage[T1]{fontenc}    % Zeichen-Enkodierung
\usepackage{blindtext}      % Platzhaltertexte
\usepackage{amsmath}        % Mathematische Umgebungen und Symbole
\usepackage{algorithm}      % Umgebung für Algorithmen
\usepackage[noend]{algpseudocode} % Pseudocode innerhalb der Algorithmen
\usepackage{graphicx}       % Einbinden von Grafiken
\usepackage{xcolor}         % Farbnamen und Definition
\usepackage{subcaption}     % Subcaption für Unterabbildungen
\usepackage{array}
\usepackage{geometry}
\usepackage{hyperref} % For clickable URLs in bibliography
\usepackage{url}      % For proper formatting of URLs
\geometry{a4paper, margin=1in}
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily,
  breaklines=true,     % Enables automatic line breaking
  linewidth=\linewidth % Limits the width to the current line width
}
\usepackage[acronym]{glossaries}
\makeglossaries
% \setlength{\columnsep}{0.5in}

% \usepackage[numbers,sort&compress]{natbib}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Eigene Farben definieren
\definecolor{twblue}{HTML}{1f497d}
\definecolor{structure}{HTML}{1f497d}
\definecolor{blindtext}{HTML}{D3D3D3}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Dokumentbeginn
\begin{document}

\title{Nonlinear Model Predictive Control-Based Local Planner with Integrated Dynamic Obstacle Avoidance for Mobile Robots}

\author{\IEEEauthorblockN{\textbf{Student:} Eppacher Kevin, BSc. \textbf{PK:} 2310331013}
        \IEEEauthorblockN{\textbf{Peer ReviewerIn:} Schwaiger Simon, MSc.}
}

@@

\begin{abstract}

% \title{nMPC-Local-Planner with Integrated Dynamic Obstacle Avoidance}

Mobile robots play a crucial role in various domains, including logistics and service industries, where they must navigate dynamic environments filled with fast-moving obstacles. Traditional hierarchical navigation systems separate global path planning and local obstacle avoidance, with global planners focusing on static maps and local planners using real-time sensor data for immediate adjustments. However, local planners such as DWA, TEB, and MPC struggle in highly dynamic environments due to issues such as slow responsiveness, difficulty in handling sharp turns or dense obstacles, and high computational demands, which can result in suboptimal paths or failure to avoid collisions in real time. This paper proposes a nonlinear Model Predictive Controller (nMPC) integrated with an obstacle detection algorithm to improve dynamic obstacle avoidance performance. The nMPC framework selectively extracts only the most relevant obstacles for avoidance, reducing the computational load while maintaining safety and efficiency. This optimization minimizes computation time, reduces cross-track errors, and ensures effective navigation. Through comprehensive simulations, the nMPC demonstrated superior performance compared to state-of-the-art planners in dynamic obstacle avoidance tasks, achieving faster arrival times and lower tracking errors. Future work will extend the application of this controller to real-world systems, enhancing the nMPC’s adaptability to various robotic platforms.
The project is made available on GitHub for public use and contributions at the following link: \url{https://github.com/KevinEppacher/walle\_ws.git}
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Mobile robots are becoming increasingly indispensable in various fields such as intralogistics, public, and private services, with applications ranging from autonomous floor-cleaning robots to vacuum cleaners. In dynamic environments, these robots are frequently confronted with fast-moving obstacles, which they must detect and avoid in real-time.

Navigation in mobile robotics is generally approached using a hierarchical paradigm, where a global planner generates a high-level path and a local planner refines this path based on real-time sensory data. The global planner operates over a static map, which is typically generated using techniques like occupancy grids, and focuses on finding an optimal or feasible path from the start to the goal location. This path is often generated by algorithms such as Dijkstra, A*, or more advanced techniques like Rapidly Exploring Random Trees (RRT) \cite{thrun2005probabilistic}.

However, the global path alone is insufficient in dynamic environments where obstacles may move or change over time. To handle these situations, the local planner operates at a higher frequency, continuously refining the robot's trajectory to avoid newly detected obstacles, while ensuring the robot still follows the general direction of the global path. The local planner relies on real-time sensory input (e.g., LiDAR or camera data) to make instantaneous adjustments, balancing both collision avoidance and adherence to the global plan \cite{khatib1986real, siegwart2011autonomous}.

This separation between global path planning, which is concerned with long-term planning over a static map, and local obstacle avoidance, which focuses on immediate reactive control, is key to ensuring that the robot can navigate effectively in both known and dynamic environments. By maintaining this hierarchical structure, robots are able to achieve long-distance navigation goals while remaining responsive to dynamic obstacles in real-time \cite{khatib1986real, siegwart2011autonomous}.

@@

\cite{ASAP} conducted a comprehensive comparison of its developed local planner, ASAP, with other seminal planners, including DWA, TEB, and MPC. In \cite{ASAP}, we compare the performance of these planners in guiding a mobile robot to follow a global path while avoiding densely placed obstacles. In this study, \cite{ASAP} evaluated the local planners based on metrics such as arrival time, cross track error, and computation time.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1\columnwidth]{images/DWA_TEB_MPC_ASAP_stats.png}
    \caption{Simulation results for arrival time, cross track error and computation time for DWA, TEB, MPC and ASAP (directly taken from \cite{ASAP})}
    \label{fig:DWA_TEB_MPC_ASAP_stats}
\end{figure}

@@

Similarly, \cite{MPCvsTEB} performed a comparison between MPC and TEB planners. The study evaluated both the TEB planner and various forms of MPC in terms of arrival time, path length, control effort, and CPU time. The results indicated that the TEB planner and the quadratic form of MPC had comparable arrival times, around 126.9 seconds. However, the time-optimal MPC, which is specifically optimized for minimizing arrival time, completed the task in just 116 seconds.

% Problem description arising from motivation (was fehlt im aktuellen sota -> vermeidung dynamischer Objekte)
Based on the findings in \cite{ASAP}, the weaknesses of the local planners DWA, TEB and MPC are highlighted, particularly in the context of dynamic obstacle avoidance.

DWA faces significant challenges when obstacles are directly in front of the robot, especially in environments with narrow passages or sharp turns. This is due to the fact that DWA samples a wide range of speeds to find a feasible collision-free trajectory, but this method increases computation time and is prone to failure in tight spaces. In addition, in environments with dynamic obstacles, velocity sampling often cannot react quickly enough, making DWA less reliable in real-time applications \cite{ASAP}.

TEB, known for generating smooth and collision-free paths, works by solving a non-convex optimisation problem. While TEB performs well in static or moderately dynamic environments, its performance degrades in highly dynamic environments due to the need for frequent re-optimisation. This can lead to suboptimal paths or deadlocks where the planner cannot find a valid solution, especially when obstacles move rapidly or unpredictably. In addition, TEB's computation time increases significantly in dense environments, further limiting its real-time applicability \cite{ASAP, MPCvsTEB}.

MPC, on the other hand, provides near-optimal trajectories by predicting future robot states and optimising control inputs based on system dynamics and constraints. However, solving these constrained optimisation problems at each time step imposes a high computational burden, especially when a longer prediction horizon is required to handle complex or dynamic environments. The real-time performance of MPC is highly dependent on the available computational resources and the complexity of the environment. Furthermore, the effectiveness of MPC is highly dependent on an accurate model of the robot and the environment, which can be difficult to achieve in real-world scenarios with unpredictable obstacles \cite{ASAP, MPCvsTEB}.

% Paper contribution
The main objective is to design a controller that effectively addresses the challenges of dynamic obstacle avoidance, offering improved performance in real-time navigation within dynamic environments. This work contributes to the State-of-the-art development, of a nonlinear Model Predictive Controller for a mobile robot with integrated dynamic sparse obstacle avoidance. This work proposed an nMPC local planner, in which the nMPC is integrated with an obstacle avoidance algorithm, to reduce computation time and increase performance, which is evaluated by reducing the arrival time and the cross-tracking error.

@@

This section provides an overview of several state-of-the-art local controllers, including the Dynamic Window Approach (DWA), Timed Elastic Band (TEB) and Model Predictive Control (MPC). In contrast, nonlinear Model Predictive Control (nMPC) is presented to overcome these limitations and provide improved performance.

\subsection{Preliminary Information on Optimization-Based Methods}

Optimization-based methods like DWA, TEB, and MPC solve an Optimal Control Problem (OCP) by minimizing a cost function while respecting system constraints. The cost function is an objective measure that typically balances factors such as trajectory tracking, control effort, and obstacle avoidance. Its goal is to find control inputs that minimize deviation from the desired path while maintaining safety and efficiency.

In OCPs, equality constraints ensure the system dynamics are followed, while inequality constraints handle limits such as collision avoidance and control input bounds. The problem is often discretized using methods like multiple shooting or single shooting, which transform the continuous problem into a finite-dimensional optimization problem.

These optimization problems are solved in real-time using solvers such as Quadratic Programming (QP), Interior Point Optimizer (IPOPT), or Sequential Quadratic Programming (SQP), which can handle both the constraints and optimization efficiently over the prediction horizon.

\subsection{Dynamic Window Approach (DWA)}
DWA is a popular local planner that selects safe and feasible velocities by sampling translational \((v)\) and rotational \((\omega)\) velocities within a dynamic window, reinitialising the current pose and velocities and the current obstacle position. These sampled velocities are evaluated through an objective function that balances forward movement and obstacle avoidance. DWA computes a set of trajectories and selects the one that maximizes progress toward the goal while ensuring the robot avoids collisions \cite{DWA, fox1997dynamic}.

The velocities \(v\) and \(\omega\) are constrained within bounds (see Eq. \ref{eq:dwa_constraints}):

@@

The objective function \(G(v, \omega)\) evaluates the alignment with the goal, distance to obstacles, and forward velocity (see Eq. \ref{eq:dwa_objective}):

@@

While DWA is efficient and handles real-time obstacle avoidance well, it struggles with dynamic obstacles due to its lack of future obstacle movement prediction. The extensive velocity sampling can lead to increased computation times in complex environments \cite{DWA}.

@@

\subsection{Timed Elastic Band (TEB)}
TEB models the robot’s trajectory as a dynamic band that adjusts to avoid obstacles while satisfying kinodynamic constraints. The trajectory is continuously optimized using a cost function that takes into account the goal, collisions, and system dynamics \cite{inproceedings}.

@@

\begin{equation}
\label{eq:teb_cost}
\begin{aligned}
J(\mathbf{x}, \mathbf{v}) = \sum_{i=1}^{n-1} \big( & \text{cost}_{\text{goal}}(x_i) + \text{cost}_{\text{collision}}(x_i) \\
& + \text{cost}_{\text{kinodynamics}}(v_i) \big)
\end{aligned}
\end{equation}

@@

Recent improvements to TEB incorporate dynamic obstacle velocities using Kalman filtering to predict obstacle movement and optimize the trajectory in real time. This improved TEB method \cite{pr12050984} enhances real-time performance by reducing the overall computational burden while maintaining a high success rate for dynamic obstacle avoidance.

\subsection{Model Predictive Control (MPC)}

Model Predictive Control (MPC) computes optimal control inputs over a prediction horizon by solving an optimization problem that accounts for both system dynamics and environmental constraints. The fundamental principle of MPC is to generate a trajectory by forward-simulating the system’s dynamics over a prediction horizon, while adhering to constraints and minimizing a cost function. At each time step, the optimization problem is solved, and only the first control input is applied, making the process iterative and reactive to dynamic environments.

The optimization problem involves the system model \(x_{k+1} = f(x_k, u_k)\), where \(x_k\) represents the system state and \(u_k\) represents the control input at time step \(k\). The MPC aims to minimize the cost function \(J(\vec{x}_k, \vec{u}_k)\), as shown in Equation \eqref{eq:mpc_cost}:

\begin{equation}
\label{eq:mpc_cost}
\min_{\vec{u}_k} J(\vec{x}_k, \vec{u}_k) = \phi(\vec{x}(N)) + \sum_{k=0}^{N-1} L(\vec{x}(k), \vec{u}(k))
\end{equation}

where \( \phi(\vec{x}(N)) \) is the terminal cost, and \(L(\vec{x}(k), \vec{u}(k))\) represents the running cost over each time step \(k\). The term \( \phi(\vec{x}(N)) \) penalizes the deviation of the final state \(\vec{x}(N)\) from the desired target state, ensuring that the robot closely follows the planned trajectory by the end of the prediction horizon. The running cost \(L(\vec{x}(k), \vec{u}(k))\) is a quadratic function of both the state and the control input, where:

- \( Q \) is the state weighting matrix, which penalizes deviations of the robot’s state from the desired reference trajectory. Larger values in \( Q \) correspond to higher penalties for errors in specific state variables, such as position or velocity.

- \( R \) is the control input weighting matrix, which penalizes large or sudden control inputs. This encourages smoother and more efficient control actions, avoiding erratic behavior.

- \( S \) is the terminal cost weighting matrix, which emphasizes the importance of reaching the target state at the end of the prediction horizon, ensuring that the final state is as close as possible to the desired state. The overall objective of the cost function is to balance minimizing the tracking error while ensuring efficient use of control inputs.

Compared to DWA and TEB, MPC offers superior handling of dynamic obstacles by continuously adjusting the robot’s trajectory in real-time, optimizing for both safety and performance. Unlike DWA, which samples a range of velocities and often struggles with sharp turns or dense obstacle fields due to its reactive nature, MPC employs a predictive approach that anticipates future states, making it more adaptable to dynamic scenarios \cite{ASAP}. Additionally, while TEB excels at generating smooth paths, it frequently encounters computational challenges in complex environments, especially when faced with multiple moving obstacles. The non-convex optimization problem that TEB solves can result in suboptimal paths or deadlocks, particularly when obstacles move unpredictably \cite{ASAP, MPCvsTEB}.

MPC, on the other hand, directly integrates the system’s nonlinear dynamics and constraints into its optimization process. By predicting future states and control inputs over a finite prediction horizon, MPC is capable of generating near-optimal paths that ensure collision-free navigation, even in highly dynamic environments. An additional advantage of the nMPC controller is its ability to enforce both hard and soft constraints. The soft constraints allow for minor violations (penalizing the cost function when predicted states approach obstacles), thereby guiding the robot towards safer paths. The hard constraints, on the other hand, guarantee a strictly collision-free trajectory by setting firm limits that cannot be breached. This dual-constraint structure provides flexibility in optimization while ensuring safety \cite{wan2024MultiUAVsNMPC}.

However, one of the long-standing challenges of MPC has been its computational intensity. Solving a constrained optimization problem in real-time, especially with a large prediction horizon, demands significant computational resources \cite{MPC_ROS, Model_predictive_Controller_for_Mobile_Robot}. To overcome this challenge, a sparse obstacle representation is integrated into the proposed nMPC framework. By selectively focusing on the most relevant obstacles and discarding less critical data, the computational load is reduced, allowing for real-time performance while retaining the predictive capabilities of MPC. This integration not only lowers the computational cost but also enhances the planner’s ability to efficiently handle dynamic environments with high obstacle density \cite{Lindqvist2020MPC}.

@@

Figure \ref{fig:nMPC_Local_Planner_Diagram_2} illustrates the individual components of the nMPC local planner. The main elements are the trajectory planner, nMPC and the obstacle detection algorithm.

The trajectory planner receives the global path from a planner based on Dijkstra's search. It processes the global plan into a local plan, taking into account the number of predictions and the prediction horizon length specified by the nMPC. After the nMPC receives the reference trajectory, the robot’s current pose, and the target State, the optimizer of the nMPC calculates the optimal control signal \(\mathbf{u} = \begin{bmatrix} v & \omega \end{bmatrix}^\top\), until the target is reached.

The robot is localised in a predefined map using probabilistic localisation methods \cite{thrun2005probabilistic}. Based on the scan data from the LiDAR, the DDMR can detect obstacles. The Obstacle Detection Node filters the LiDAR data to define a search cone, reducing unnecessary computational load on the nMPC. It then sends the positions of the obstacles, along with the predefined safety radius of the obstacles, to the trajectory planner, which passes this information to the nMPC for further processing. By sending only the most relevant obstacles to the nMPC, the computational load is minimized, allowing for real-time decision-making.

@@

\subsection{Nonlinear Model Predictive Controller}

The nMPC is responsible for planning a path based on the current pose, the target pose, and the reference trajectory, using an optimizer that not only satisfies the kinematics of a differential drive mobile robot (DDMR) but also avoids obstacles and respects limitations on control inputs and States. Figure \ref{fig:nMPC_Blockdiagram_2} illustrates the components of an nMPC \cite{falcone2007predictive}.

\begin{figure}[!h]
    \centering
    \includegraphics[width=1\columnwidth]{images/nMPC_Blockdiagram_2.png}
    \caption{Detailed Blockdiagram of nMPC (modified from \cite{Effects_of_roll_dynamics_for_car_stability_control_by_laguerre_functions})}
    \label{fig:nMPC_Blockdiagram_2}
\end{figure}

First, the Optimal Control Problem (OCP) must be converted into a Nonlinear Programming Problem (NLP) using the multiple shooting discretization method. Unlike single shooting, where only the control vector is optimized and the entire system is forward-simulated, multiple shooting introduces both the control vector and the state vector as optimization variables. This means that not only are the control inputs \( \mathbf{u} \) optimized, but the system states \( \mathbf{x} \) at each prediction step are also included in the optimization \cite{BOCK19841603}.

The cost function (see Eq. \ref{eq:L_function}) is not merely a summation over all time steps, but instead, each segment's dynamics must adhere to the system's model equations, expressed as equality constraints (see Fig. \ref{fig:nMPC_Blockdiagram_2}). This enforces that the model dynamics are satisfied over the entire prediction horizon, and discrepancies between predicted and simulated states are minimized, improving both stability and accuracy \cite{DIEHL2002577}.

Moreover, this approach significantly enhances the computational time efficiency. By breaking the problem into smaller segments, the optimization solver can converge faster and more reliably, especially in the presence of complex nonlinear dynamics or large prediction horizons. The ability to treat state and control variables as separate optimization entities helps in avoiding divergence \cite{BOCK19841603}.

The goal is to minimize the cost function (see Eq. \ref{eq:L_function}), which consists of the States and control actions for all predictions \cite{rawlings2009model}.

\begin{equation}
\label{eq:L_function}
\begin{aligned}
\min_{\vec{u}_k} J(\vec{x}_k, \vec{u}_k) = J_{\text{obs}} + \phi(\vec{x}_N) 
+ &\\ \sum_{k}^{k+N-1}
(\vec{x}_{R,k} - \vec{x}_{\text{ref},k})^T Q (\vec{x}_{R,k} - \vec{x}_{\text{ref},k})&+ \vec{u}_k^T R \vec{u}_k
\end{aligned}
\end{equation}

The cost function \(J\) consists of three main components:
\begin{itemize}
    \item \(\phi(\vec{x}_N)\) represents the terminal cost, ensuring the final State \(\vec{x}_N\) is close to the target State, as defined in Equation \ref{eq:phi_definition} \cite{rawlings2009model}.
    \begin{equation}
    \label{eq:phi_definition}
    \phi(\vec{x}(N)) = \frac{1}{2} \vec{x}_N^T S \vec{x}_N
    \end{equation}

    \item \(J_{\text{obs}}\) is a penalty term that increases the cost if the robot comes close to an obstacle, calculated according to Equation \ref{eq:J_obs} \cite{frasch2013parallel}.
\end{itemize}

\begin{equation}
\label{eq:J_obs}
J_{\text{obs}} = \sum_{k=0}^{N} \sum_{i=1}^{n_{\text{obs}}} \frac{w_{\text{pen}}}{\sqrt{(x_k - x_{\text{obs},i})^2 + (y_k - y_{\text{obs},i})^2} + \epsilon}
\end{equation}

where \(x_{\text{obs},i}\) and \(y_{\text{obs},i}\) represent the coordinates of the \(i\)-th obstacle, and \(\epsilon\) is a small value to avoid singularities. 
The State vector \(\mathbf{x}\) and control vector \(\mathbf{u}\) are defined in Equations \ref{eq:State_vector} and \ref{eq:control_vector}, respectively \cite{liniger2015nonlinear}.

\begin{equation}\label{eq:State_vector}
    \mathbf{x}_k = \begin{bmatrix} x_k & y_k & \theta_k \end{bmatrix}^T
\end{equation}

\begin{equation}\label{eq:control_vector}
    u_k = \begin{bmatrix}
    v_k & \omega_k
\end{bmatrix}^T
\end{equation}

These vectors describe the position, orientation, and control inputs of the robot at each time step \(k\). The evolution of the State is constrained to follow the robot's kinematics, as enforced by Equation \ref{eq:discrete_dynamics} \cite{liniger2015nonlinear}.
